%============================= abs.tex================================
%============================= abs.tex================================
\begin{Abstract}

Polynomial commitment schemes (PCS) are a core building block of
modern zero-knowledge proof systems, yet there is little empirical
data comparing different constructions under identical hardware
conditions across multiple elliptic curves.

This thesis benchmarks four PCS: KZG, Multilinear KZG (MKZG),
Bulletproofs, and Dory. KZG, MKZG, and Bulletproofs are implemented
from scratch in Rust using \texttt{arkworks}; Dory uses the a16z
reference implementation. All four are evaluated on BLS12-381, BN254,
and BLS12-377 across seven metrics: setup, commit, prove, verify,
scalar multiplication, pairing, and proof size. The effect of
Pippenger's multi-scalar multiplication (MSM) algorithm is also
studied empirically.

KZG achieves the smallest proof (48\,bytes) and fastest verification
(3.8\,ms, constant in $n$), making it optimal when a trusted setup is
acceptable. MKZG extends KZG to multilinear polynomials but incurs
exponential setup cost in the number of variables $\mu$ and an
$O(\log n)$ verifier (21.85\,ms at $n=1024$). Dory delivers
$O(\log n)$ verification (95.8\,ms at $n=1024$) with no trusted
setup, at the cost of larger proofs ($\approx$20\,KB). Bulletproofs
also need no trusted setup but suffer from $O(n)$ verification,
limiting practical use. Across all schemes, BN254 is approximately
$1.9\times$ faster than BLS12-381, while BLS12-377 is comparable to
BLS12-381. Pippenger's algorithm gives a 5--6$\times$ speedup on
large single MSMs.

\end{Abstract}
%=======================================================================

%=======================================================================
